---
title: "Exercise 3"
author: "Oleg Kartavtsev"
date: "5/24/2022"
output: github_document
---

### Part 1: Importing the datasets, loading the libraries and adding "race", "gender" and "tenure" columns
```{r}

library(arrow)
library(readr)
applications <- read_parquet(paste0('C:\\Users\\oleg1\\Desktop\\McGill\\Classes\\Organizational Network Analysis\\app_data_sample.parquet'))
edges <- read_csv(paste0('C:\\Users\\oleg1\\Desktop\\McGill\\Classes\\Organizational Network Analysis\\edges_sample.csv'))

library(gender)
library(tidyverse)
#install_genderdata_package() # only run this line the first time you use the package, to get data for it

# get a list of first names without repetitions
examiner_names <- applications %>% 
  distinct(examiner_name_first)


# get a table of names and gender
examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )


# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()

library(wru)

examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()


examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()


examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))



# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)

applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()

library(lubridate) # to work with dates

examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 


examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))

examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
  ) %>% 
  filter(year(latest_date)<2018)

applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()

```
### Part 2: choosing to workgroups to focus on (in my case, 163 and 172) and printing out their summaries, as well as plotting the distribution of gender and race across these workgroups
``` {r}
#we pick 2 workgroups to focus on: 163 and 172

#creating a subset and summaries of the workgroup 163
wg163 <- subset(applications, grepl("^163", applications$examiner_art_unit))
wg163$gender <- factor(wg163$gender)
wg163$race <- factor(wg163$race)
summary(wg163$gender)
summary(wg163$race)
summary(wg163$tenure_days)

```

```{r}
#creating a subset and summaries of the workgroup 172
wg172 <- subset(applications, grepl("^172", applications$examiner_art_unit))
wg172$gender <- factor(wg172$gender)
wg172$race <- factor(wg172$race)
summary(wg172$gender)
summary(wg172$race)
summary(wg172$tenure_days)
```

```{r}
#combining 2 datasets together (for plotting purposes)
library(dplyr)
wg163$workgroup <- c('wg_163')
wg172$workgroup <- c('wg_172')
combined <- union(x=wg163,y=wg172)
```

```{r}
#plot gender and race across the workgroups
library(ggplot2)
ggplot(combined, aes(workgroup, ..count..)) + geom_bar(aes(fill = race), position = "dodge")
ggplot(combined, aes(workgroup, ..count..)) + geom_bar(aes(fill = gender), position = "dodge")

toPlot<-combined%>%
  group_by(workgroup, gender, race)%>%
  summarise(n = n())%>%
  group_by(race)%>%
  mutate(prop = n/sum(n))


ggplot(data = toPlot, aes(workgroup, prop, fill = gender)) + 
  geom_col() + 
  facet_grid(~race)+
  scale_fill_manual(values = c("white","green", "red", "blue"))
```

Results of the analysis: workgroup 172 is significantly male-dominated while workgroup 163 has only a slightly higher male distribution. In addition, workers in both groups are predominantly White, with the Asian workers having significance presence as well. Upon a closer look, it seems that all Black workers in group 163 are female and practically all Hispanic workers in both groups (more so in group 163) are male. The male-female ratio is the closest to the equal among White workers in the workgroup 163.

```{r}
#setting up the variables and tables
library(igraph)
library(tidygraph)
library(ggraph)
#creating a dataset with distinct information about only people from workgroups of our interest (in this case, 163 and 172) using both ego and alter nodes
distinct_groups = distinct(subset(applications, select=c(examiner_art_unit, examiner_id)))
distinct_groups$wg = substr(distinct_groups$examiner_art_unit, 1,3)
distinct_groups = distinct_groups[distinct_groups$wg==163 | distinct_groups$wg==172,]
merged = merge(x=edges, y=distinct_groups, by.x="ego_examiner_id", by.y="examiner_id", all.x=TRUE)
merged = merged %>% rename(ego_art_unit=examiner_art_unit, ego_wg=wg)
merged = drop_na(merged)
merged = merge(x=merged, y=distinct_groups, by.x="alter_examiner_id", by.y="examiner_id", all.x=TRUE)
merged = merged %>% rename(alter_art_unit=examiner_art_unit, alter_wg=wg)
merged = drop_na(merged)
#finalizing the nodes table
New_nodes = subset(merged, select=c(ego_examiner_id,ego_art_unit, ego_wg)) %>% rename(examiner_id=ego_examiner_id,art_unit=ego_art_unit,wg=ego_wg)
alterNodes = subset(merged, select=c(alter_examiner_id,alter_art_unit, alter_wg))%>% rename(examiner_id=alter_examiner_id,art_unit=alter_art_unit,wg=alter_wg)
nodes = rbind(New_nodes, alterNodes)
nodes = distinct(nodes)
nodes = nodes %>% group_by(examiner_id) %>% summarise(examiner_id=first(examiner_id), art_unit=first(art_unit), wg=first(wg))

```


```{r}
#plotting the directed graph
graph1 = graph_from_data_frame(d=merged, vertices=nodes, directed=TRUE)

V(graph1)$workgroup = nodes$art_unit

```

```{r}
Degree <- degree(graph1, v=V(graph1))
V(graph1)$size = Degree
ggraph(graph1, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size=size, color=workgroup), show.legend=T)
```

From this visualization (based on the degree centrality), it seems that there isn't much overlap between the two randomly chosen groups (except for a handful of individual cases and a single outlier in one of the larger sections of the network). Overall impression that this graph creates is that people prefer to ask for an advice within their respective workgroups.

```{r}
e <- evcent(graph1)$vector
V(graph1)$evcent <- round(e,2)

plot(graph1, layout=layout.fruchterman.reingold,
    vertex.size = 6,          
    vertex.label = NA,
    vertex.label.cex = 0.8,   
    vertex.label.dist = 1.5,  
    vertex.label.color = "black",
    vertex.color = graph1$workgroup)
```

Eigenvector is similar to degree centrality, but the difference is that eigenvector also takes nodes' power into account. This graph is different from the previous one in the sense that there is much more overlap between different groups, with clusters of homogenous workgroups (denoted by identical colors) being present, but barely to the same extent. This leads to the conclusion that, when we take into account nodes' relative power, there is a notable overlap between the workgroups. 

```{r}
degree(graph1)
```

```{r}
betweenness(graph1, directed = TRUE)
```

```{r}
closeness(graph1)
```

Bonus: if our goal is not to look at the big picture but at individual workers and their impact on the workplace, it would be more convenient to calculate raw centrality scores, as it makes it easier to identify distinct employer_ids. For example, according to the degree centrality, employee with id "94782" is directly connected to 50 other employees (nodes)! That's quite impressive. Knowing their employer ID, we can search it in the main "applications" dataset to get more information regarding them.
